---
title: "CRE Biobanking Project Update"
author: "Laura L. Cowden"
date: "2024-12-05"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include = FALSE}
library(tidyverse)
library(janitor)
library(readxl)
library(lubridate)
library(dplyr)

setwd("H:/AMT/CRE Biobanking 2024")

# load data sets
carba <- read_csv("Carbapenemase Detection.csv") %>%
  clean_names()
all <- read_csv("CRE_biobanking_all.csv") %>%
  clean_names()
```

```{r carbapenemases detected, include = FALSE}
# separate date into separate columns for month, day and year

carba <- carba %>%
  separate(., "date_collected", c("month", "day", "year"), sep = "/")

# statistics
carba %>%
  group_by(carbapenemase_detected, year) %>%
  summarize(count = n())
```

```{r all data, include = FALSE}
# create new dataset of frequencies per last name
all <- all %>%
  separate(., "collect_date", c("month", "day", "year"), sep = "/")

patient_freq <- all %>%
  group_by(last_name) %>% 
  summarize(count=n())

patient_freq %>%
  count(count)
```

<font size="5"> CRE Biobanking Data Review </font> 

As of December 5th 2024, a total of 733 CRE isolates have been collected spanning years 2018-2024.

There was a notable drop in collections in recent years; potentially partly due to the influx of new micro techs at the AMT bench not putting them aside. Re-training occurred in summer of 2024 in attempt to correct this.

```{r totals per year, echo=FALSE}
ggplot(all, aes(year))+
  geom_bar(fill = "cyan4")+
  xlab("number of CRE isolates collected")
```

<font size="4"> Opportunities for Longitudinal Study </font> 

The following histogram plots the number of isolates collected per any one patient over time:

```{r long, echo=FALSE}
ggplot(patient_freq, aes(count))+
  geom_bar(fill = "cornflowerblue")+
  xlab("number of CRE isolates collected per any one patient")+
  scale_x_continuous(labels = 0:23, breaks = 0:23) 
  
```

<font size="4"> Carbapenemases Detected </font> 

Clinical microbiology began testing for the top-5 carbapenemase enzymes in late 2022. These data consist of isolates tested between that time and November 2024.

```{r source, echo=FALSE}
# carbapenemases detected by source
ggplot(carba, aes(carbapenemase_detected, fill = source))+
         geom_bar()+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5))+
  labs(title = "Carbapenemases detected by Source")
```

```{r species, echo=FALSE}
# carbapenemases detected by species
ggplot(carba, aes(carbapenemase_detected, fill = species))+
  geom_bar()+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5))+
  labs(title = "Carbapenemases Detected by Species")
```

```{r year, echo=FALSE}
# carbapenemases detected by year
ggplot(carba, aes(carbapenemase_detected, fill = year))+
  geom_bar()+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5))+
  labs(title = "Carbapenemases Detected by Year")
```

